### 2016-05-02 submission-petter.tex
Additions to experimental setup and data

%
% File acl2016.tex
%
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

%% Greit å vite:

%% Siteringer:
%% Med parentes rundt hele: \cite{Gusfield:97}
%% Eller med parentes kun rundt år:  \newcite{Gusfield:97}
%% Flere på en gang: \cite{Gusfield:97,Aho:72}

%% Fra instruksjonene, om \aclfinalcopy :

%% By uncommenting {\small\verb|\aclfinalcopy|} at the top of this document, it
%% will compile to produce an example of the camera-ready formatting; by leaving
%% it commented out, the document will be anonymized for initial submission.  When
%% you first create your submission on softconf, please fill in your submitted
%% paper ID where {\small\verb|***|} appears in the
%% {\small\verb|\def\aclpaperid{***}|} definition at the top.

%% The review process is double-blind, so do not include any author information
%% (names, addresses) when submitting a paper for review.  However, you should
%% maintain space for names and addresses so that they will fit in the final
%% (accepted) version.  The ACL 2016 \LaTeX\ style will create a titlebox space of
%% 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

%% \subsection{The Ruler}
%% The ACL 2016 style defines a printed ruler which should be presented in the
%% version submitted for review.  The ruler is provided in order that
%% reviewers may comment on particular lines in the paper without
%% circumlocution.  If you are preparing a document without the provided
%% style files, please arrange for an equivalent ruler to
%% appear on the final output pages.  The presence or absence of the ruler
%% should not change the appearance of any other content on the page.  The
%% camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
%% the {\small\verb|\aclfinalcopy|} command in the document preamble.)

\title{The greatest PoS tags, just the best really}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain}  \\\And
  Third Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Abstraction ensues.
\end{abstract}

%\section{Credits}

\section{Introduction}
\label{sec:intro}

\section{Previous/Related work}
\label{sec:prev}

Megyesi, Swedish

MacKinlay, English

OBT, Norwegian

\section{Data}
\label{sec:data}
\subsection{The Norwegian Dependency Treebank}
The newly developed Norwegian Dependency Treebank (NDT) \cite{Sol:Skj:Ovr:14}
is the first publicly available treebank for Norwegian. The treebank was used
for training and evaluation in our experiments. The tag set consists of 12
morphosyntactic PoS tags, with 7 additional tags for punctuation and symbols.

\paragraph{PoS Tag Set}
\paragraph{Dependency Relations}

\section{Experimental Setup}
\label{sec:setup}

\paragraph{Tagger}
TnT \cite{Bra:00}

\paragraph{Parser}
In choosing a syntactic parser for our experiments, we considered previous work
on dependency parsing of Norwegian, specifically that of \cite{Sol:Skj:Ovr:14}.
They found the Mate parser \cite{Boh:10} to be the most successful parser for
the parsing of NDT. Recent dependency parser comparisons \cite{Cho:Tet:Ste:15}
also showed that Mate performed very well on parsing of the English portion of
the OntoNotes 5 corpus, beating a range of contemporary state-of-the-art
parsers.

\paragraph{Tag Set Mapping}
In order to alter the tag set of NDT, we created a mapping for carrying out the
tag set modifications. We created a mapping for carrying out the tag set
modifications that maps the relevant existing tags to new, more fine-grained
tags including more relevant morphological features for the applicable tokens.

\paragraph{Baseline}
It is common practice to compare the performance of PoS taggers to a
pre-computed baseline for an initial point of comparison.
For PoS tagging, a commonly used baseline is the Most Frequent Tag (MFT)
baseline, which we use in our experiments. This involves labeling each word
with the tag it was assigned most frequently in the training. All unknown
words, i.e., words not seen in the training data, are assigned the tag most
frequently assigned to words seen only once in the training. Unknown and
infrequent words have in common that they rarely occur, and we might therefore
expect them to have similar properties.
%, there are two main
%approaches: assign it the most frequent tag overall in the training data, or
%the tag most frequently assigned to words seen only once in the training data.
%We adopted the latter approach for our baseline tagger, as we believe that
%unknown words may have much in common with words that occur only once in the
%training, more so than simply words of the most frequent part-of-speech. The
%reason for this is that unknown and infrequent words have in common that they
%rarely occur, and we might therefore expect them to have similar properties.

\paragraph{Tags \& Features}
As we seek to quantify the effects of PoS tagging in a realistic setting, we
want to run the parser on automatically assigned PoS tags. For the training of
the parser, however, we have two options: using either gold standard or
automatically assigned tags. In order to settle on a configuration, we
conducted experiments with gold standard and automatically assigned tags to see
how they differ with respect to performance. The results of our experiments
reveal that the combination of training and testing on automatic tags is
superior to training on gold standard tags and testing on automatic tags,
surprisingly. Consequently, the parser was both trained and tested on
automatically assigned tags in our experiments.

Note that it is absolutely crucial that the morphological features in the
treebank are removed when using automatic tags, as they are still gold
standard. For instance, if a verb token is erroneously tagged as a noun, we
could potentially have a noun token with verbal features such as tense, which
markedly obfuscates the training and parsing. Another important factor is that
we want to isolate the effect of PoS tags, necessitating the exclusion of
morphological features.

\section{Tag Set Optimization}
\label{sec:optimization}

%\begin{table}
    %\centering
    %%\smaller[0.5]
    %\begin{tabular}{@{}ll@{}}
        %\toprule
        %\textbf{Tag} & \textbf{Description} \\
        %\midrule
        %\texttt{adj|komp} & Comparative adjective \\
        %\texttt{adj|pos} & Positive adjective \\
        %\texttt{adj|sup} & Superlative adjective \\
        %\texttt{det|be} & Definite determiner \\
        %\texttt{det|ub} & Indefinite determiner \\
        %\texttt{pron|pers} & Personal pronoun \\
        %\texttt{pron|pers|akk} & Personal pronoun, accusative \\
        %\texttt{pron|pers|nom} & Personal pronoun, nominative \\
        %\texttt{pron|refl} & Reflexive pronoun \\
        %\texttt{pron|res} & Reciprocal pronoun \\
        %\texttt{pron|sp} & Interrogative pronoun \\
        %\texttt{subst|appell} & Common noun \\
        %\texttt{subst|appell|be} & Common noun, definite \\
        %\texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
        %\texttt{subst|appell|ub} & Common noun, indefinite \\
        %\texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
        %\texttt{subst|prop} & Proper noun \\
        %\texttt{subst|prop|gen} & Proper noun, genitive \\
        %\texttt{verb|fin} & Finite verb \\
        %\texttt{verb|infin} & Nonfinite verb \\
        %\bottomrule
    %\end{tabular}
    %\caption{The optimized tag set.}
    %\label{optimizedtagset}
%\end{table}

\section{Optimized Pipeline}
\label{sec:pipeline}

\section{Summary/Conclusion and Future Work}
\label{sec:summary}

% \section*{Acknowledgments}

\bibliographystyle{acl2016}
\bibliography{references}
% \appendix
%\section{Supplemental Material}
%\label{sec:supplemental}

\end{document}