### 2016-05-02 submission-petter.tex
Minor additions to optimization section

%
% File acl2016.tex
%
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{acl2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

%% Greit å vite:

%% Siteringer:
%% Med parentes rundt hele: \cite{Gusfield:97}
%% Eller med parentes kun rundt år:  \newcite{Gusfield:97}
%% Flere på en gang: \cite{Gusfield:97,Aho:72}

%% Fra instruksjonene, om \aclfinalcopy :

%% By uncommenting {\small\verb|\aclfinalcopy|} at the top of this document, it
%% will compile to produce an example of the camera-ready formatting; by leaving
%% it commented out, the document will be anonymized for initial submission.  When
%% you first create your submission on softconf, please fill in your submitted
%% paper ID where {\small\verb|***|} appears in the
%% {\small\verb|\def\aclpaperid{***}|} definition at the top.

%% The review process is double-blind, so do not include any author information
%% (names, addresses) when submitting a paper for review.  However, you should
%% maintain space for names and addresses so that they will fit in the final
%% (accepted) version.  The ACL 2016 \LaTeX\ style will create a titlebox space of
%% 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

%% \subsection{The Ruler}
%% The ACL 2016 style defines a printed ruler which should be presented in the
%% version submitted for review.  The ruler is provided in order that
%% reviewers may comment on particular lines in the paper without
%% circumlocution.  If you are preparing a document without the provided
%% style files, please arrange for an equivalent ruler to
%% appear on the final output pages.  The presence or absence of the ruler
%% should not change the appearance of any other content on the page.  The
%% camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
%% the {\small\verb|\aclfinalcopy|} command in the document preamble.)

\title{The greatest PoS tags, just the best really}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain}  \\\And
  Third Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Abstraction ensues.
\end{abstract}

%\section{Credits}

\section{Introduction}
\label{sec:intro}

\section{Previous/Related work}
\label{sec:prev}
Previous work related to our experiments with tag set modifications include
investigating the effects of PoS tag sets on tagging of Swedish
\cite{Meg:01,Meg:02,Meg:09} and English \cite{Mac:05}. The important difference
is that they investigate the effects of PoS tags on tagging, while we assess
its effects on a downstream application, namely syntactic parsing.

\section{Data}
\label{sec:data}
\subsection{The Norwegian Dependency Treebank}
We used the newly developed Norwegian Dependency Treebank (NDT)
\cite{Sol:Skj:Ovr:14}, which is the first publicly available treebank for
Norwegian (for our experiments). It was developed at the National Library of
Norway in collaboration with the University of Oslo, and contains manual
syntactic and morphological annotation. The treebank contains 311 000 tokens
of Bokmål and 303 000 tokens of Nynorsk, the morphological annotation of which
follows that of the Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in
turn is largely based on the work of \newcite{Faa:Lie:Van:97}. The annotated
texts are mostly newspaper text, but also include government reports,
parliament transcripts and excerpts from blogs. The annotation process of the
treebank was supported by the Oslo-Bergen Tagger and then manually corrected by
annotators.

Noe om OBT? hmmm

\paragraph{PoS Tag Set}
The tag set consists of 12 morphosyntactic PoS tags, with 7 additional tags for
punctuation and symbols. The tag set is thus rather coarse-grained, with broad
categories such as \texttt{subst} (noun) and \texttt{verb} (verb).

\paragraph{Dependency Relations}

\section{Experimental Setup}
\label{sec:setup}
In preparation to conducting our experiments with linguistically motivated tag
set modifications, a concrete setup for the experiments needed to be
established, which is presented in the following.

\paragraph{Data Set Split}
As there was no standardized data set split of NDT due to its very recent
development, we needed to establish a data set split
(training/development/test) in preparation to our experiments. Our data set
split of the treebank follows the standard 80-10-10 (training/development/test)
split and will be distributed with the treebank and proposed as the new
standard(?). In creating the data set split, care has been taken to preserve
contiguous texts in the various data sets while keeping the split balanced in
terms of genre (and source). Our proposed data set split was used in the
Norwegian contribution to the Universal Dependencies project \cite{Ovr:Hoh:16}.
The split will be made available at a companion website.

\paragraph{Tagger}
For our experiments with tag set modifications, we wanted a PoS tagger that is
both fast and accurate. There is often a trade-off between the two, as the best
taggers tend to suffer in terms of speed due to their complexity. However, a
tagger that achieves both close to state-of-the-art accuracy as well as very
high speed is TnT \cite{Bra:00}. The fact that TnT was used for evaluating the
universal tag set \cite{Pet:Das:McD:12}, served as another good indication of
TnT being appropriate for our task. The sum of these factors led to TnT being
the tagger of choice for our experiments.

\paragraph{Parser}
In choosing a syntactic parser for our experiments, we considered previous work
on dependency parsing of Norwegian, specifically that of \cite{Sol:Skj:Ovr:14}.
They found the Mate parser \cite{Boh:10} to be the most successful parser for
the parsing of NDT. Furthermore, recent dependency parser comparisons
\cite{Cho:Tet:Ste:15} showed that Mate performed very well on parsing of the
English portion of the OntoNotes 5 corpus, beating a range of contemporary
state-of-the-art parsers.

\paragraph{Tag Set Mapping}
In order to alter the tag set of NDT, we created a mapping for carrying out the
tag set modifications. We created a mapping for carrying out the tag set
modifications that maps the relevant existing tags to new, more fine-grained
tags including more relevant morphological features for the applicable tokens.

\paragraph{Baseline}
It is common practice to compare the performance of PoS taggers to a
pre-computed baseline for an initial point of comparison.
For PoS tagging, a commonly used baseline is the Most Frequent Tag (MFT)
baseline, which we use in our experiments. This involves labeling each word
with the tag it was assigned most frequently in the training. All unknown
words, i.e., words not seen in the training data, are assigned the tag most
frequently assigned to words seen only once in the training. Unknown and
infrequent words have in common that they rarely occur, and we might therefore
expect them to have similar properties.
%, there are two main
%approaches: assign it the most frequent tag overall in the training data, or
%the tag most frequently assigned to words seen only once in the training data.
%We adopted the latter approach for our baseline tagger, as we believe that
%unknown words may have much in common with words that occur only once in the
%training, more so than simply words of the most frequent part-of-speech. The
%reason for this is that unknown and infrequent words have in common that they
%rarely occur, and we might therefore expect them to have similar properties.

\paragraph{Tags \& Features}
As we seek to quantify the effects of PoS tagging in a realistic setting, we
want to run the parser on automatically assigned PoS tags. For the training of
the parser, however, we have two options: using either gold standard or
automatically assigned tags. In order to settle on a configuration, we
conducted experiments with gold standard and automatically assigned tags to see
how they differ with respect to performance. The results of our experiments
reveal that the combination of training and testing on automatic tags is
superior to training on gold standard tags and testing on automatic tags,
surprisingly. Consequently, the parser was both trained and tested on
automatically assigned tags in our experiments.

Note that it is absolutely crucial that the morphological features in the
treebank are removed when using automatic tags, as they are still gold
standard. For instance, if a verb token is erroneously tagged as a noun, we
could potentially have a noun token with verbal features such as tense, which
markedly obfuscates the training and parsing. Another important factor is that
we want to isolate the effect of PoS tags, necessitating the exclusion of
morphological features.

\section{Tag Set Optimization}
\label{sec:optimization}

For each tag, we first experiment with each of the features in isolation before
employing various combinations of them. We base our choices of combinations on
how promising the features are and what we deem worth investigating in terms of
linguistic utility, in order to see how the features might interact.

%\begin{table}
    %\centering
    %%\smaller[0.5]
    %\begin{tabular}{@{}ll@{}}
        %\toprule
        %\textbf{Tag} & \textbf{Description} \\
        %\midrule
        %\texttt{adj|komp} & Comparative adjective \\
        %\texttt{adj|pos} & Positive adjective \\
        %\texttt{adj|sup} & Superlative adjective \\
        %\texttt{det|be} & Definite determiner \\
        %\texttt{det|ub} & Indefinite determiner \\
        %\texttt{pron|pers} & Personal pronoun \\
        %\texttt{pron|pers|akk} & Personal pronoun, accusative \\
        %\texttt{pron|pers|nom} & Personal pronoun, nominative \\
        %\texttt{pron|refl} & Reflexive pronoun \\
        %\texttt{pron|res} & Reciprocal pronoun \\
        %\texttt{pron|sp} & Interrogative pronoun \\
        %\texttt{subst|appell} & Common noun \\
        %\texttt{subst|appell|be} & Common noun, definite \\
        %\texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
        %\texttt{subst|appell|ub} & Common noun, indefinite \\
        %\texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
        %\texttt{subst|prop} & Proper noun \\
        %\texttt{subst|prop|gen} & Proper noun, genitive \\
        %\texttt{verb|fin} & Finite verb \\
        %\texttt{verb|infin} & Nonfinite verb \\
        %\bottomrule
    %\end{tabular}
    %\caption{The optimized tag set.}
    %\label{optimizedtagset}
%\end{table}

\section{Optimized Pipeline}
\label{sec:pipeline}

\section{Summary/Conclusion and Future Work}
\label{sec:summary}

% \section*{Acknowledgments}

\bibliographystyle{acl2016}
\bibliography{references}
% \appendix
%\section{Supplemental Material}
%\label{sec:supplemental}

\end{document}