### 2016-05-05 submission.tex
Minor changes, fixed references

%
% File acl2016.tex
%
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{acl2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}
\usepackage{relsize}
\usepackage{multirow}
%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

%% Greit å vite:

%% Siteringer:
%% Med parentes rundt hele: \cite{Gusfield:97}
%% Eller med parentes kun rundt år:  \newcite{Gusfield:97}
%% Flere på en gang: \cite{Gusfield:97,Aho:72}

%% Fra instruksjonene, om \aclfinalcopy :

%% By uncommenting {\small\verb|\aclfinalcopy|} at the top of this document, it
%% will compile to produce an example of the camera-ready formatting; by leaving
%% it commented out, the document will be anonymized for initial submission.  When
%% you first create your submission on softconf, please fill in your submitted
%% paper ID where {\small\verb|***|} appears in the
%% {\small\verb|\def\aclpaperid{***}|} definition at the top.

%% The review process is double-blind, so do not include any author information
%% (names, addresses) when submitting a paper for review.  However, you should
%% maintain space for names and addresses so that they will fit in the final
%% (accepted) version.  The ACL 2016 \LaTeX\ style will create a titlebox space of
%% 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

%% \subsection{The Ruler}
%% The ACL 2016 style defines a printed ruler which should be presented in the
%% version submitted for review.  The ruler is provided in order that
%% reviewers may comment on particular lines in the paper without
%% circumlocution.  If you are preparing a document without the provided
%% style files, please arrange for an equivalent ruler to
%% appear on the final output pages.  The presence or absence of the ruler
%% should not change the appearance of any other content on the page.  The
%% camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
%% the {\small\verb|\aclfinalcopy|} command in the document preamble.)

\title{Optimizing a PoS Tag Set for Dependency Parsing}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain}  \\\And
  Third Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Abstraction ensues.
\end{abstract}

%\section{Credits}

\section{Introduction}
\label{sec:intro}
Part-of-speech (PoS) tagging is an important preprocessing step for many NLP
tasks, such as dependency parsing \cite{Niv:Hal:Kub:07,Haj:Cia:Joh:09}, named
entity recognition \cite{Tjo:DeM:03} and sentiment analysis
\cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the development of
PoS-taggers -- to the effect that this task is often considered more or less a
solved task -- considerably less effort has been devoted to the empirical
evaluation of the PoS tag sets themselves. Error analysis of PoS taggers
indicate that, whereas tagging improvement through means of learning algorithm
or feature engineering seems to have reached something of a plateau, linguistic
and empirical assessment of the distinctions made in the PoS tag sets may be a
venue worth investigating further \cite{Man:11}. Clearly, the utility of a PoS
tag set is tightly coupled with the downstream task for which it is performed.
Even so, PoS tag sets are usually employed in a ``one size fits all'' fashion,
regardless of the requirements posed by the task which makes use of this
information.

It is well known that syntactic parsing often benefits from quite fine-grained
morphological distinctions \cite{Zha:Niv:11,See:Kuh:13}. Morphology interacts
with syntax though phenomena such as agreement and case marking, and
incorporating information on morphological properties of words can therefore
often improved parsing performance. However, in a realistic setting where the
aim is to automatically parse raw text, the generation of morphological
information will often require a separate step of morphological analysis that
can be quite costly.

In this paper, we optimize a PoS tag set for the task of data-driven dependency
parsing of Norwegian. We report on a set of experiments where various
morphological distinctions are introduced to the PoS annotations and evaluated
both intrinsically, i.e., in terms of PoS tagging accuracy, and extrinsically,
in terms of parsing accuracy. Our results show that the introduction of
morphological distinctions not present in the original tag set, whilst
compromising tagger accuracy, actually leads to significantly improved parsing
accuracy. This optimization allows us to bypass the additional step of
morphological analysis, framing the whole preprocessing problem as a simple
tagging task.

The article is structured as follows. We start out by reviewing previous work
on tag set evaluation in Section \ref{sec:prev}, while Section \ref{sec:data}
details the treebank that provides the initial gold annotations used for our
experiments. Section \ref{sec:setup} describes the experimental setup used in
our work and Section \ref{sec:optimization} goes on to provide the results from
our tag set optimization. Finally, Section \ref{sec:summary} summarizes our
main findings and discusses some avenues for future work.

\section{Previous Work}
\label{sec:prev}
There has been little previous work dedicated specifically to
extrinsic evaluation of the effects of PoS tag sets on downstream
applications.  There has, however, been some work that evaluates the
effects of PoS tag set granularity on PoS tagging. This includes
investigation of the effects of PoS tag sets on tagging of Swedish
\cite{Meg:01,Meg:02} and English \cite{Mac:05}.

%Increases in tagger accuracy does not readily translate to
%improved parsing.

\newcite{Meg:02} trained and evaluated a range of PoS taggers on the
Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, they
investigated the effects of tag set size on tagging by mapping the original tag
set into smaller subsets designed for parsing. They argue that a tag set with
complete morphological tags may not be necessary for all NLP applications, for
instance syntactic parsing. They found that the smallest tag set comprising 26
tags yields the lowest tagger error rate.  However, for some of the taggers,
augmenting the tag set with more linguistically informative tags may actually
lead to a drop in error rate. They argue that this shows that the size of the
tag set as well as the type of information in the tags are crucial factors for
tagger performance.  However/unfortunately, they do not report results of
parsing with the various PoS tag sets.

Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
performance in English, specifically the Wall Street Journal portion of Penn
Treebank \cite{Mar:San:Mar:93}. They mapped the original tag set of Penn
Treebank comprising 45 tags to more fine-grained tag sets using linguistic
insight to investigate whether additional linguistic information included in
finer-grained tags could assist the tagger. Experimenting with both lexically
and syntactically conditioned modifications such as distinguishing between
count nouns and noncount nouns and between transitive and intransitive verbs,
they found that more fine-grained tag sets rarely led to improvements in tagger
accuracy; the most successful modification yielded an improvement in tagger
accuracy of 0.05 percentage points.  They did not find any statistically
significant improvements using linguistically informed distinctions, arguing
that their results do not support the hypothesis that it is possible to achieve
significant performance improvements in PoS tagging by utilizing a
finer-grained tag set.

Transition/overgang her ...

\section{The Norwegian Dependency Treebank}
\label{sec:data}
We used the newly developed Norwegian Dependency Treebank (NDT)
\cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
Norwegian. It was developed at the National Library of Norway in
collaboration with the University of Oslo, and contains manual
syntactic and morphological annotation. The treebank contains data
from both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
tokens of Nynorsk. We will in the following only be using the Bokmål
portion of the treebank. The annotated texts are mostly newspaper
text, but also include government reports, parliament transcripts and
excerpts from blogs. The annotation process of the treebank was supported by
the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually
corrected by annotators, adding syntactic dependency analysis to the
morphosyntactic annotation.

\paragraph{Morphological Annotation}
The morphological annotation and PoS tag set of NDT follows the Oslo-Bergen
Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work
of \newcite{Faa:Lie:Van:97}. The tag set consists of 12 morphosyntactic PoS
tags outlined in Table \ref{ndttagset}, with 7 additional tags for punctuation
and symbols. The tag set is thus rather coarse-grained, with broad categories
such as \texttt{subst} (noun) and \texttt{verb} (verb). The PoS tags are
complemented by a large set of morphological features, providing information
about morphological properties such as definiteness, number and tense. These
features are used in our tag set modifications, where the coarse PoS tag of
relevant tokens is concatenated with one or more of these features to include
more linguistic information in the tags.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Tag} & \textbf{Description} \\
        \midrule
        \texttt{adj} & Adjective \\
        \texttt{adv} & Adverb \\
        \texttt{det} & Determiner\\
        \texttt{inf-merke} & Infinitive marker \\
        \texttt{interj} & Interjection \\
        \texttt{konj} & Conjunction \\
        \texttt{prep} & Preposition \\
        \texttt{pron} & Pronoun \\
        \texttt{sbu} & Subordinate conjunction \\
        \texttt{subst} & Noun \\
        \texttt{ukjent} & Unknown (foreign word) \\
        \texttt{verb} & Verb \\
        \bottomrule
    \end{tabular}
    \caption{Overview of the PoS tag set of NDT.}
    \label{ndttagset}
\end{table}

\paragraph{Syntactic Annotation}
The syntactic annotation choices in NDT are largely based on the Norwegian
Reference Grammar \cite{Faa:Lie:Van:97}. The annotation choices are outlined in
Table \ref{ndtannotation}, taken from/courtesy of \newcite{Sol:Skj:Ovr:14},
providing overview of the analyses of syntactic constructions that often
distinguish dependency treebanks, such as coordination and the treatment of
auxiliary and main verbs. The set of dependency relations comprises 29
dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
(subject) and \textsc{koord} (coordination).

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Head} & \textbf{Dependent} \\
        \midrule
        Preposition & Prepositional complement \\
        Finite verb & Complementizer \\
        First conjunct & Subsequent conjuncts \\
        Finite auxiliary & Lexical/main verb \\
        Noun & Determiner \\
        \bottomrule
    \end{tabular}
    \caption{Central head-dependent annotation choices in NDT.}
    \label{ndtannotation}
\end{table}

\section{Experimental Setup}
\label{sec:setup}
In preparation to conducting our experiments with linguistically motivated tag
set modifications, a concrete setup for the experiments needed to be
established, which is presented in the following.

Our approach is somewhat semi-automatic in that we employ a "hybrid"
combination of manual annotation and automatic labeling in our experiments. Our
initial tags are gold standard, and using linguistic insight coupled with
computational considerations, we introduce new features or feature combination
to the tag set. These tags are run through a pipeline with automatically
assigned tags in both training and testing, where the most promising
modifications are used "further". Hence, the tag set modifications are not
deterministic, nor do we try all possible combination; we only experiment with
linguistically motivated distinctions. We are in reality altering the gold
standard tags by the addition of more linguistic information.

\paragraph{Data Set Split}
As there was no standardized data set split of NDT due to its very recent
development, we needed to establish a data set split
(training/development/test). Our data set split of the treebank follows the
standard 80-10-10 (training/development/test) split and will be distributed
with the treebank and proposed as the new standard(?). In creating the data set
split, care has been taken to preserve contiguous texts in the various data
sets while keeping the split balanced in terms of genre. Our
proposed data set split was used in the Norwegian contribution to the Universal
Dependencies project \cite{Ovr:Hoh:16}. The split will be made available at a
companion website. In our experiments, we train on the training data and
evaluate on the development data, before finally evaluating our optimized tag
set on the held-out test data set.

\paragraph{Tagger}
For our experiments with tag set modifications, we sought a PoS tagger both
reasonably fast and accurate. There is a often a considerable trade-off between
the two factors, as the most accurate taggers tend to suffer in terms of speed
due to their complexity. However, a tagger that achieves both close to
state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. TnT
was furthermore recently used to evaluate the recently proposed universal tag
set \cite{Pet:Das:McD:12}. The sum of these factors led to TnT being the tagger
of choice for our experiments.

\paragraph{Parser}
In choosing a syntactic parser for our experiments, we considered previous work
on dependency parsing of Norwegian, specifically that of
\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to be the
most successful parser for parsing of NDT. Furthermore, recent dependency
parser comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well
on parsing of English, beating a range of contemporary state-of-the-art
parsers. Mate was consequently selected to evaluate the tag set modifications in
our experiments.

\paragraph{Tag Set Mapping}
The tag set modifications are realized by mapping the existing tags to new,
more fine-grained tags "based" on associated morphological features to be
included in specified tags. The procedure maps the relevant existing tags to
new, more fine-grained tags including more relevant morphological features for
the applicable tokens, i.e., tokens that are assigned said tag and feature(s)
in the gold standard data. Hence, we are in actuality altering the gold
standard tags by the addition of more relevant linguistic information,
producing new gold standard annotations of higher quality.

%This is done by concatenating the original tag with the
%specified features for all applicable tokens

%contains the set of features in its morphological features.

\paragraph{Evaluation}
To evaluate tagging and parsing with the various tag set modifications in our
experiments, we employed commonly used evaluation metrics. Tagging is evaluated
in terms of accuracy (reported by the TnT-included \texttt{tnt-diff} script),
while labeled attachment score (LAS) and unlabeled attachment score (UAS) are
computed by the
\texttt{eval.pl}\footnote{\url{http://ilk.uvt.nl/conll/software.html}} script
(used in the CoNLL shared tasks) to evaluate the parsing.

\paragraph{Baseline}
It is common practice to compare the performance of PoS taggers to a
pre-computed baseline for an initial point of comparison.  For PoS tagging, a
commonly used baseline is the Most Frequent Tag (MFT) baseline, which we used
in our experiments. This involves labeling each word with the tag it was
assigned most frequently in the training. All unknown words, i.e., words not
seen in the training data, are assigned the tag most frequently assigned to
words seen only once in the training. Unknown and infrequent words have in
common that they rarely occur, and we might therefore expect them to have
similar properties.

\paragraph{Tags \& Features}
As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
in application to raw text, we want to evaluate the parser on automatically
assigned PoS tags. For the training of the parser, however, we have two
options: using either gold standard or automatically assigned tags. In order to
settle on a configuration, we conducted experiments with gold standard and
automatically assigned tags to see how they differ with respect to performance.
The results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
the combination of training and testing on automatic tags is clearly superior
to training on gold standard tags and testing on automatic tags. Consequently,
the parser was both trained and tested on automatically assigned tags in our
experiments.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Training} & \textbf{Testing} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        Gold & Gold & 90.15\% & 92.51\% \\
        Gold & Auto & 85.68\% & 88.98\% \\
        Auto & Auto & 87.01\% & 90.19\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of parsing with Mate using the various tag configurations,
        either using gold standard tags or automatically assigned from TnT.
        \emph{Gold} denotes gold standard tags, \emph{Auto} denotes
        automatically assigned tags from TnT.}
    \label{tagconfigeval}
\end{table}

We removed any morphological features in order to simulate a realistic setting.
Moreover, it is crucial that we remove these features when working with
automatically assigned tags, as the features are still gold standard. For
instance, if a verb token is erroneously tagged as a noun, we could potentially
have a noun token with verbal features such as tense, which markedly obfuscates
the training and parsing. Another important aspect is that we want to isolate
the effect of PoS tags, necessitating the exclusion of morphological features.

In the following, we report on a set of experiments which evaluate increasingly
fine-grained PoS tag sets. The PoS tag sets are motivated by morphosyntactic
properties of Norwegian and are evaluated in terms of both tagger and parser
accuracy.

\section{Tag Set Optimization}
\label{sec:optimization}
By introducing more fine-grained linguistically motivated distinctions in a tag
set, we increase the linguistic information represented in the tags, which may
assist the parser in recognizing and generalizing syntactic patterns. However,
the addition of more linguistic information to the tags and thus a more
fine-grained tag set will most likely lead to a drop in tagger accuracy due to
the increase in complexity. The best tagging does not necessarily lead to the
best parse, hence it is interesting to investigate how linguistically informed
tag set modifications may affect the interplay between tagging and parsing.
%As
%we want to investigate how we can improve the linguistic quality of a tag set,
%we will
%only consider tag set modifications we deem linguistically sensible.
%we will not introduce tag set modifications which are not linguistically
%motivated. We will only consider distinctions we deem linguistically sensible.

\subsection{Baseline Experiments}
In an initial round of experiments, we concatenated the tag of each token with
its set of morphological features in order to map the original tag set to a
new, more fine-grained tag set (hereafter referred to as the \emph{full} tag
set). This resulted in a total of 368 tags, which is clearly very fine-grained.
The two initial tag sets, i.e., the original tag set comprising 19 tags and the
full tag set comprising 368 tags, thus represent two extremes in terms of
granularity. To evaluate the initial tag sets and investigate how the tag set
granularity would affect the performance of tagging and parsing, we ran
baseline experiments with the two tag sets. In Table \ref{inittagseteval}, we
report the results of these experiments. We see that the tagger accuracy
drastically drops when going from the original to the full tag set. TnT reports
an accuracy of 97.47\% on the original tag set, which is reduced to 93.48\% for
the full tag set. These results confirm our hypothesis that the very high
linguistic quality in the full, fine-grained tag set comes at the expense of
drops in tagger performance.  However, the additional linguistic information
provided by the full tag set improves the parser performance. With the original
tag set, Mate reports an LAS of 87.01\% and a UAS of 90.19\%, which increases
to 87.15\% and 90.39\%, respectively, using the full tag set. As we are looking
for linguistically informed distinctions that improve the syntactic parsing,
these results are promising and serve to indicate that additional linguistic
information assists the syntactic parsers, motivating the optimization of the
existing PoS tag set.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}lllll@{}}
        \toprule
        \textbf{Tag set} & \textbf{MFT} & \textbf{Accuracy} &
        \textbf{LAS} & \textbf{UAS} \\
        \midrule
        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the two initial tag sets.}
    \label{inittagseteval}
\end{table}

\subsection{Tag Set Experiments}
We modified the tags for nouns (\texttt{subst}), verbs (\texttt{verb}),
adjectives (\texttt{adj}), determiners (\texttt{det}) and pronouns
(\texttt{pron}) in NDT by appending selected sets of morphological features to
each tag in order to increase the linguistic information expressed by the tags.
For each tag, we first experimented with each of the available features in
isolation before employing various combinations of them. We based our choices
of combinations on how promising the features are and what we deem worth
investigating in terms of linguistic utility, in order to see how the features
might interact.

The morphological properties of the various parts-of-speech are reflected in
the morphological features associated with the respective PoS tags. For
instance, as nouns in Norwegian can take on gender, definiteness and number,
additionally marked with genitive case, the treebank operates with features for
gender, definiteness, number and case.
%with "accompanying" values
In addition to morphological properties such as definiteness, tense and number,
all classes except for verbs can be distinguished "on" type, e.g., common nouns
and proper nouns, while pronouns can be reflexive, reciprocal, personal or
interrogative.

%"Nearly" all tag set modifications leads to a drop in tagger accuracy, as
%expected. Few exceptions... We observed increases in parser accuracy scores
%with linguistically motivated tag set modifications for all PoS tags we
%experimented with, the most improved being noun with an increase in LAS from
%the original tag set by 1.80 percentage points.

\paragraph{Nouns}
In Norwegian, nouns are assigned gender (feminine, masculine or neuter),
definiteness (definite or indefinite) and number (singular or plural). There
is agreement in gender, definiteness and number between nouns and their
modifiers (adjectives and determiners). Additionally, NDT has a separate case
feature for distinguishing nouns in genitive case.

As nouns constitute the largest class in the treebank by far, the effects are
greatest for noun tokens in terms of overall change in performance. The results
of experimenting with tag set modifications for nouns are shown in Table
\ref{substresults}, and reveal that, apart from case, none of the tag set
modifications improves the tagging. However, they all give rise to increases in
parser accuracy scores. Genitive case marks possession, hence nouns marked with
genitive case are quite different from other nouns, taking a noun phrase as
complement. Distinguishing on type is useful and informative, as evident by the
presence of separate tags for proper and common nouns in many tag sets, such as
those of Penn Treebank and Stockholm-Umeå corpus.

The most informative features are definiteness, which
leads to an increase in LAS by 1.26 percentage points to 88.27\%, and type,
yielding an LAS of 88.07\%. Turning to combinations of features, we found that
the combination of type and case, as well as type and definiteness, were the
most promising, which led us to combine type, case and definiteness in a final
experiment, resulting in LAS of 88.81\% and UAS of 91.73\%, constituting large
increases from parsing with the original tag set, by 1.80 percentage points
and 1.54 percentage points, respectively.

\begin{table}
    \centering
    \smaller[1]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS}
        \\
        \midrule
        --- & 97.47\% & 87.01\% & 90.19\% \\
        Case & \textbf{97.48\%} & 87.63\% & 90.72\% \\
        Definiteness & 97.00\% & 88.27\% & 91.42\% \\
        Gender & 96.09\% & 87.21\% & 90.36\% \\
        Number & 96.37\% & 87.97\% & 91.00\% \\
        Type & 96.92\% & 88.07\% & 91.11\% \\
        Case \& definiteness & 97.03\% & 88.39\% & 91.44\% \\
        Type \& case & 96.92\% & 88.46\% & 91.51\% \\
        Type \& definiteness & 96.99\% & 88.44\% & 91.48\% \\
        Type \& number & 96.37\% & 87.95\% & 90.95\% \\
        Type, case \& definiteness & 97.05\% & \textbf{88.81\%} &
        \textbf{91.73\%} \\
        %Type, definiteness \& number & 96.46\% & 88.06\% & 91.12\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for nouns.}
    \label{substresults}
\end{table}

\paragraph{Verbs}
Verbs are inflected for tense (infinitive, present, preterite, past perfect) in
Norwegian and additionally exhibit mood (imperative, indicative) and voice
(active or passive). Note that both voice and mood have only a single value in
the treebank, \texttt{pass} (passive) and \texttt{imp} (imperative),
respectively.  Verbs which are not passive are implicitly active, and verbs
which are not imperative are indicative.

Table \ref{verbresults} presents the results from tagging and parsing with
modified verb tags. Introducing features in isolation, mood is the only feature
leading to an increase in LAS, with a reported LAS of 87.04\%. Imperative
clauses are fundamentally different from indicative clauses, as they lack an
overt subject.  Although voice is a very distinguishing property for verbs, and
passive verbs appear in clauses very different from those with active verbs,
introducing this distinction in the tag set leads to a drop in LAS of 0.05
percentage points.  Distinguishing between the various tenses yields an LAS of
86.97\%. Combining the two most promising features, mood and tense, resulted in
an LAS of 87.12\% and UAS of 90.31\%.

In an additional experiment, we mapped the verb tenses (and mood, in the case
of imperative) to finiteness. All verbs have finiteness, hence this distinction
has broad coverage. This mapping is syntactically grounded as finite verbs and
nonfinite verbs appear in completely different syntactic constructions, and
proved to greatly improve the parsing, as we saw the overall largest parser
accuracy scores, with 87.30\% for LAS and 90.43\% for UAS, 0.29 and 0.24
percentage points higher than the baseline, respectively. This coincides with
the observations seen for Swedish in \newcite{Ovr:08}, where finiteness was
found to be a very beneficial linguistic feature for parsing.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        Mood & 97.43\% & 87.04\% & 90.19\% \\
        Tense & 97.30\% & 86.97\% & 90.18\% \\
        Voice & 97.45\% & 86.96\% & 90.09\% \\
        Mood \& tense & 97.31\% & 87.12\% & 90.31\% \\
        Voice \& tense & 97.28\% & 86.99\% & 90.15\% \\
        Mood, tense \& voice & 97.27\% & 86.83\% & 90.05\% \\
        Finiteness & 97.35\% & \textbf{87.30\%} & \textbf{90.43\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for verbs.}
    \label{verbresults}
\end{table}

\paragraph{Adjectives}
Adjectives agree with the noun they modify in terms of gender, number and
definiteness in Norwegian. Adjectives are also inflected for degree (positive,
comparative or superlative).

Table \ref{adjresults}

All features except for number led to increases in parser accuracy scores.
Degree being the most promising with a reported LAS of 87.29\%, while
distinguishing adjectives on definiteness yield an LAS of 87.14\% and gender
leads to LAS of 87.10\%.

Combining the most promising features, we found that none of the combinations
surpass the LAS with the distinction of degree alone. Definiteness and degree
with an LAS of 87.23\% and definiteness and number with an LAS of 87.27\% and
UAS identical to that of degree alone.

Turning to combinations of features, definiteness and number achieve the best
results, very close to that of degree, with 0.02 percentage points lower LAS
and identical UAS. Adjectives agree with their head noun and determiner in
definiteness and number, making this an expected improvement. The combination
of definiteness and degree is also quite promising, obtaining LAS of 87.23\%
and UAS of 90.39\%. It is interesting that none of the combinations surpass the
experiment with degree alone, which indicates that degree does not interact
with the other features in any syntactically significant way.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS}
        \\
        \midrule
        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        Definiteness & 96.84\% & 87.14\% & 90.29\% \\
        Degree & 97.41\% & \textbf{87.29\%} & \textbf{90.44\%} \\
        Gender & 96.89\% & 87.10\% & 90.25\% \\
        Number & 96.71\% & 86.99\% & 90.10\% \\
        Type & 97.40\% & 87.11\% & 90.25\% \\
        Definiteness \& degree & 96.81\% & 87.23\% & 90.39\% \\
        Definiteness \& gender & 96.31\% & 87.18\% & 90.39\% \\
        Definiteness \& number & 96.78\% & 87.27\% & \textbf{90.44\%} \\
        Degree \& gender & 96.87\% & 87.00\% & 90.16\% \\
        Degree \& number & 96.76\% & 87.13\% & 90.26\% \\
        %Definiteness, degree \& number & 96.81\% & 87.14\% & 90.30\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for adjectives.}
    \label{adjresults}
\end{table}

\paragraph{Determiners}
Like adjectives, determiners in Norwegian agree with the noun they modify in
terms of gender, number and definiteness. Possessive pronouns are
treated/analyzed as determiners in NDT.

For a number of reasons, we do not report results from combinations of features
for determiners, as the features could not be combined in any meaningful way.
For instance, determiners in plural never have marked definiteness, hence this
combination would only apply to determiners in singular and roughly correspond
to the distinction on definiteness alone. Moreover, determiners in plural are
not distinguished in terms of gender (thus ruling out the combination of number
and gender), neither are definite determiners. Another aspect taken into
consideration is that the most promising distinction, definiteness, applies to
such a small number of tokens that more fine-grained distinctions would be
overly sparse.

The results from the experiments with determiners are shown in Table
\ref{detresults}.  Introducing the distinction of type led to an increase in
tagger accuracy by 0.14 percentage points to 97.61\%, while marginally
impacting the parsing, with LAS of 87.00\%, 0.01 percentage points below that
of the original tag set.  The increase in tagger accuracy when introducing the
distinction of type is noteworthy, as we expected the finer granularity to lead
to a decrease in accuracy. This serves to indicate that more fine-grained
distinctions for determiners, which is a quite disparate category in the
treebank, may be quite useful for tagging. However, as it has negative impact
on the syntactic parsing, we can conclude that the type of a determiner does
not assist in generalizing syntactic patterns, as most determiners, regardless
of type, appear in the same syntactic constructions (i.e., before a noun or
noun phrase).

Gender, on the other hand, improved the parsing, but complicated the tagging,
as the various genders are often difficult to differentiate, especially so in
the case of masculine and feminine, which share many of the same determiners.
The number of a determiner, i.e., singular or plural, led to a small increase
in tagger accuracy and LAS, while marginally lower UAS, 90.18\%, 0.01
percentage points lower than that of the original tag set. The
introduction of definiteness to the determiners led to the best parsing
results, LAS of 87.30\% and UAS of 90.42\%, while also increasing the tagger
accuracy slightly. The increase in LAS and UAS is rather interesting, as there
are only 121 "determiner tokens" with marked definiteness in the development
data. As this accounts for a very small number of tokens, coupled with the
previously noted considerations, we did not consider further fine-grained
modifications with definiteness. This goes to show that tokens with overt
definiteness have noticeable impact on the syntactic parsing, and that
distinguishing on definiteness is very beneficial.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        --- & 97.47\% & 87.01\% & 90.19\% \\
        Definiteness & 97.49\% & \textbf{87.30\%} & \textbf{90.42\%} \\
        Gender & 97.28\% & 87.09\% & 90.31\% \\
        Number & 97.49\% & 87.04\% & 90.18\% \\
        Type & \textbf{97.61\%} & 87.00\% & 90.11\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for determiners.}
    \label{detresults}
\end{table}

\paragraph{Pronouns}
Pronouns in Norwegian include personal, reciprocal, reflexive and
interrogative. They can exhibit gender, number and person. Personal pronouns
have case (accusative or nominative).

The results in Table \ref{pronresults} show that number, person and type are
the most informative features for parsing, with LAS of 87.21\%, 87.22\% and
87.19\%, respectively. However, when combining number and person, we observe a
drop by more than 0.2 percentage points, indicating that these features do not
interact in any syntactically distinctive way. The most interesting observation
is that all experiments exceed the baseline tagger accuracy, the most improved
being the most fine-grained distinction, namely type, case and number combined,
obtaining a tagger accuracy of 97.52\%. This shows that the introduction of
more fine-grained distinctions for pronouns is beneficial and aids the PoS
tagger in disambiguating ambiguous words. While case alone yields an LAS of
87.08\%, we found that the combination of type and case, which is the most
successful experiment in terms of parser performance, yields the second highest
tagging accuracy of 97.51\%. The reason for this is that pronouns of different
type and personal pronouns of different case exhibit quite different properties
and appear in different constructions. Pronouns in nominative case (i.e.,
subjects) primarily occur before the main verb, while pronouns in accusative
case (i.e., objects) occur after the main verb, as Norwegian exhibits so-called
V2 word order, requiring that the finite verb of a declarative clause appears
in the second position, hence its name. The combination of type and number
comes in close to the performance of type and case, with LAS of 87.27\% and UAS
identical to that of type and case.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        --- & 97.47\% & 87.01\% & 90.19\% \\
        Case & 97.50\% & 87.08\% & 90.21\% \\
        Gender & 97.48\% & 87.06\% & 90.23\% \\
        Number & 97.49\% & 87.21\% & 90.33\% \\
        Person & 97.49\% & 87.22\% & 90.32\% \\
        Type & 97.48\% & 87.19\% & 90.40\% \\
        Number \& person & 97.49\% & 96.98\% & 90.16\% \\
        Type \& case & 97.51\% & \textbf{87.30\%} & \textbf{90.41\%} \\
        Type \& number & 97.49\% & 87.27\% & \textbf{90.41\%} \\
        Type \& person & 97.49\% & 87.00\% & 90.14\% \\
        Type, case \& number & \textbf{97.52\%} & 87.11\% & 90.36\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for pronouns.}
    \label{pronresults}
\end{table}

\subsection{Optimized Tag Set}
The most successful tag set modifications and their results are seen in Table
\ref{respectiveresults}.

\begin{table*}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llllll@{}}
        \toprule
        \textbf{Category} & \textbf{Feature(s)} & \textbf{MFT} &
        \textbf{Accuracy} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        \emph{Baseline} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
        91.73\% \\
        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the most successful tag set
        modification for each category.}
    \label{respectiveresults}
\end{table*}

An overview of the tags comprising the final, optimized tag set can be seen in
Table \ref{optimizedtagset}.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Tag} & \textbf{Description} \\
        \midrule
        \texttt{adj|komp} & Comparative adjective \\
        \texttt{adj|pos} & Positive adjective \\
        \texttt{adj|sup} & Superlative adjective \\
        \texttt{det|be} & Definite determiner \\
        \texttt{det|ub} & Indefinite determiner \\
        \texttt{pron|pers} & Personal pronoun \\
        \texttt{pron|pers|akk} & Personal pronoun, accusative \\
        \texttt{pron|pers|nom} & Personal pronoun, nominative \\
        \texttt{pron|refl} & Reflexive pronoun \\
        \texttt{pron|res} & Reciprocal pronoun \\
        \texttt{pron|sp} & Interrogative pronoun \\
        \texttt{subst|appell} & Common noun \\
        \texttt{subst|appell|be} & Common noun, definite \\
        \texttt{subst|appell|be|gen} & Common noun, definite, genitive \\
        \texttt{subst|appell|ub} & Common noun, indefinite \\
        \texttt{subst|appell|ub|gen} & Common noun, indefinite, genitive \\
        \texttt{subst|prop} & Proper noun \\
        \texttt{subst|prop|gen} & Proper noun, genitive \\
        \texttt{verb|fin} & Finite verb \\
        \texttt{verb|infin} & Nonfinite verb \\
        \bottomrule
    \end{tabular}
    \caption{Overview of the optimized tag set.}
    \label{optimizedtagset}
\end{table}

Results from parsing with the optimized tag are shown in Table
\ref{finalresults}, compared to the results obtained with the initial tag sets.

\begin{table*}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llllll@{}}
        \toprule
        \textbf{Data Set} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Accuracy}
        & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        \multirow{2}{*}{Dev}
        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
        & Optimized & 89.20\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
        \midrule
        \multirow{2}{*}{Test}
        & Original & \textbf{94.22\%} & \textbf{97.38\%} & 86.64\% & 90.07\% \\
        %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the optimized tag set,
        compared to the initial tag sets. Trained and tested using
        automatically assigned tags from TnT.}
    \label{finalresults}
\end{table*}

%\section{Optimized Pipeline}
%\label{sec:pipeline}

\section{Summary and Future Work}
\label{sec:summary}
The improvements in parser performance with our optimized tag set indicate that
a more fine-grained PoS tag set may assist syntactic parsers in recognizing and
generalizing syntactic patterns, while potentially compromising the performance
of PoS taggers.

There are several aspects of this thesis that can be further explored in future
work, including extrinsic evaluation of the effects of PoS tag sets on other
downstream NLP applications besides parsing, such as sentiment analysis and
named entity recognition. These applications often require tagged data, but are
markedly different from syntactic parsing, hence the evaluation would involve
investigating an entirely different aspect of the effects of tag set
granularity.

% \section*{Acknowledgments}

\bibliographystyle{acl2016}
\bibliography{references}
% \appendix
%\section{Supplemental Material}
%\label{sec:supplemental}

\end{document}