### 2016-05-05 submission.tex
endringer i abstract

%
% File acl2016.tex
%
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn,
%% based on the style files for ACL-2010, which were, in turn,
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{acl2016}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}
\usepackage{booktabs}
\usepackage{relsize}
\usepackage{multirow}
%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

%% Greit å vite:

%% Siteringer:
%% Med parentes rundt hele: \cite{Gusfield:97}
%% Eller med parentes kun rundt år:  \newcite{Gusfield:97}
%% Flere på en gang: \cite{Gusfield:97,Aho:72}

%% Fra instruksjonene, om \aclfinalcopy :

%% By uncommenting {\small\verb|\aclfinalcopy|} at the top of this document, it
%% will compile to produce an example of the camera-ready formatting; by leaving
%% it commented out, the document will be anonymized for initial submission.  When
%% you first create your submission on softconf, please fill in your submitted
%% paper ID where {\small\verb|***|} appears in the
%% {\small\verb|\def\aclpaperid{***}|} definition at the top.

%% The review process is double-blind, so do not include any author information
%% (names, addresses) when submitting a paper for review.  However, you should
%% maintain space for names and addresses so that they will fit in the final
%% (accepted) version.  The ACL 2016 \LaTeX\ style will create a titlebox space of
%% 2.5in for you when {\small\verb|\aclfinalcopy|} is commented out.

%% \subsection{The Ruler}
%% The ACL 2016 style defines a printed ruler which should be presented in the
%% version submitted for review.  The ruler is provided in order that
%% reviewers may comment on particular lines in the paper without
%% circumlocution.  If you are preparing a document without the provided
%% style files, please arrange for an equivalent ruler to
%% appear on the final output pages.  The presence or absence of the ruler
%% should not change the appearance of any other content on the page.  The
%% camera ready copy should not contain a ruler. (\LaTeX\ users may uncomment
%% the {\small\verb|\aclfinalcopy|} command in the document preamble.)

\title{Optimizing a PoS Tag Set for Dependency Parsing}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain}  \\\And
  Third Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}

This paper reports on a suite of experiments that evaluates how the linguistic
granularity of part-of-speech tag sets impacts the performance of tagging and
syntactic dependency parsing. Our results show that parsing accuracy can be
significantly improved by introducing more fine-grained morphological
information in the tag set, even if tagger accuracy is compromised. Our taggers
and parsers are  trained and tested using the annotations of the Norwegian
Dependency Treebank.

\end{abstract}

%\section{Credits}

\section{Introduction}
\label{sec:intro}
Part-of-speech (PoS) tagging is an important preprocessing step for many NLP
tasks, such as dependency parsing \cite{Niv:Hal:Kub:07,Haj:Cia:Joh:09}, named
entity recognition \cite{Tjo:DeM:03} and sentiment analysis
\cite{Wil:Wie:Hof:09}. Whereas much effort has gone into the development of
PoS-taggers -- to the effect that this task is often considered more or less a
solved task -- considerably less effort has been devoted to the empirical
evaluation of the PoS tag sets themselves. Error analysis of PoS taggers
indicate that, whereas tagging improvement through means of learning algorithm
or feature engineering seems to have reached something of a plateau, linguistic
and empirical assessment of the distinctions made in the PoS tag sets may be a
venue worth investigating further \cite{Man:11}. Clearly, the utility of a PoS
tag set is tightly coupled with the downstream task for which it is performed.
Even so, PoS tag sets are usually employed in a ``one size fits all'' fashion,
regardless of the requirements posed by the task which makes use of this
information.

It is well known that syntactic parsing often benefits from quite fine-grained
morphological distinctions \cite{Zha:Niv:11,See:Kuh:13}. Morphology interacts
with syntax though phenomena such as agreement and case marking, and
incorporating information on morphological properties of words can therefore
often improved parsing performance. However, in a realistic setting where the
aim is to automatically parse raw text, the generation of morphological
information will often require a separate step of morphological analysis that
can be quite costly.

In this paper, we optimize a PoS tag set for the task of data-driven dependency
parsing of Norwegian. We report on a set of experiments where various
morphological distinctions are introduced to the PoS annotations and evaluated
both intrinsically, i.e., in terms of PoS tagging accuracy, and extrinsically,
in terms of parsing accuracy. Our results show that the introduction of
morphological distinctions not present in the original tag set, whilst
compromising tagger accuracy, actually leads to significantly improved parsing
accuracy. This optimization allows us to bypass the additional step of
morphological analysis, framing the whole preprocessing problem as a simple
tagging task.

The article is structured as follows. We start out by reviewing previous work
on tag set evaluation in Section \ref{sec:prev}, while Section \ref{sec:data}
details the treebank that provides the initial gold annotations used for our
experiments. Section \ref{sec:setup} describes the experimental setup used in
our work and Section \ref{sec:optimization} goes on to provide the results from
our tag set optimization. Finally, Section \ref{sec:summary} summarizes our
main findings and discusses some avenues for future work.

\section{Previous Work}
\label{sec:prev}
There has been little previous work dedicated specifically to
extrinsic evaluation of the effects of PoS tag sets on downstream
applications.  There has, however, been some work that evaluates the
effects of tag set granularity on PoS tagging itself. This includes
investigation of the effects of tag sets on PoS tagging for Swedish
\cite{Meg:01,Meg:02} and English \cite{Mac:05}.

%Increases in tagger accuracy does not readily translate to
%improved parsing.

\newcite{Meg:02} trained and evaluated a range of PoS taggers on the
Stockholm-Umeå Corpus (SUC) \cite{Gus:Har:06}, annotated with a tag set based
on a Swedish version of PAROLE tags totaling 139 tags. Furthermore, the effects
of tag set size on tagging was investigated by mapping the original tag
set into smaller subsets designed for parsing. \newcite{Meg:02} argues that a tag set with
complete morphological tags may not be necessary for all NLP applications, for
instance syntactic parsing. The study found that the smallest tag set comprising 26
tags yields the lowest tagger error rate.  However, for some of the taggers,
augmenting the tag set with more linguistically informative tags may actually
lead to a drop in error rate. \newcite{Meg:02} argues that this shows that the size of the
tag set as well as the type of information in the tags are crucial factors for
tagger performance.  Unfortunately, results for parsing with the various PoS
tag sets are not reported.

Similarly, \newcite{Mac:05} investigated the effects of PoS tag sets on tagger
performance in English, specifically the Wall Street Journal portion of the Penn
Treebank (PTB) \cite{Mar:San:Mar:93}. Based on linguistic
considerations \newcite{Mac:05} mapped the 45 tags of the original PTB tag set
 to more fine-grained tag sets to investigate whether additional linguistic information
could assist the tagger. Experimenting with both lexically
and syntactically conditioned modifications such as distinguishing between
count nouns and noncount nouns and between transitive and intransitive verbs,
they found that more fine-grained tag sets rarely led to improvements in tagger
accuracy; the most successful modification yielded an improvement in tagger
accuracy of 0.05 percentage points.  They did not find any statistically
significant improvements using linguistically informed distinctions, arguing
that their results do not support the hypothesis that it is possible to achieve
significant performance improvements in PoS tagging by utilizing a
finer-grained tag set.

% fixme:
Transition/overgang her ...

\section{The Norwegian Dependency Treebank}
\label{sec:data}
Our experiments our based on the newly developed Norwegian Dependency Treebank (NDT)
\cite{Sol:Skj:Ovr:14}, the first publicly available treebank for
Norwegian. It was developed at the National Library of Norway in
collaboration with the University of Oslo, and contains manually coded
syntactic and morphological annotation for
both varieties of Norwegian; 311 000 tokens of Bokmål and 303 000
tokens of Nynorsk. This paper only report results for Bokmål, the main variant.
The treebanked material mostly comprise newspaper
text, but also include government reports, parliament transcripts and
blog excerpts. The annotation process was supported by
the rule-based Oslo-Bergen Tagger \cite{Hag:Joh:Nok:00} and then manually
corrected by annotators, adding syntactic dependency analysis to the
morphosyntactic annotation.

\paragraph{Morphological Annotation}
The morphological annotation and PoS tag set of NDT is based on the same
inventory as used by the Oslo-Bergen
Tagger \cite{Hag:Joh:Nok:00,Sol:13}, which in turn is largely based on the work
of \newcite{Faa:Lie:Van:97}. The tag set consists of 12 morphosyntactic PoS
tags outlined in Table \ref{ndttagset}, with 7 additional tags for punctuation
and symbols. The tag set is thus rather coarse-grained, with broad categories
such as \texttt{subst} (noun) and \texttt{verb} (verb). The PoS tags are
complemented by a large set of morphological features, providing information
about morphological properties such as definiteness, number and tense.
Selected subsets of these
features are used in our tag set modifications, where the coarse PoS tag of
relevant tokens is concatenated with one or more features to include
more linguistic information in the tags.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Tag} & \textbf{Description} \\
        \midrule
        \texttt{adj} & Adjective \\
        \texttt{adv} & Adverb \\
        \texttt{det} & Determiner\\
        \texttt{inf-merke} & Infinitive marker \\
        \texttt{interj} & Interjection \\
        \texttt{konj} & Conjunction \\
        \texttt{prep} & Preposition \\
        \texttt{pron} & Pronoun \\
        \texttt{sbu} & Subordinate conjunction \\
        \texttt{subst} & Noun \\
        \texttt{ukjent} & Unknown (foreign word) \\
        \texttt{verb} & Verb \\
        \bottomrule
    \end{tabular}
    \caption{Overview of the original PoS tag set of NDT (excluding punctuation tags).}
    \label{ndttagset}
\end{table}

\paragraph{Syntactic Annotation}
The syntactic annotation choices in NDT are largely based on the Norwegian
Reference Grammar \cite{Faa:Lie:Van:97}. Some central annotation choices are outlined in
Table \ref{ndtannotation}, taken from of \newcite{Sol:Skj:Ovr:14},
providing overview of the analyses of syntactic constructions that often
distinguish dependency treebanks, such as coordination and the treatment of
auxiliary and main verbs. The annotations comprise 29
dependency relations, including \textsc{adv} (adverbial), \textsc{subj}
(subject) and \textsc{koord} (coordination).

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Head} & \textbf{Dependent} \\
        \midrule
        Preposition & Prepositional complement \\
        Finite verb & Complementizer \\
        First conjunct & Subsequent conjuncts \\
        Finite auxiliary & Lexical/main verb \\
        Noun & Determiner \\
        \bottomrule
    \end{tabular}
    \caption{Central head-dependent annotation choices in NDT.}
    \label{ndtannotation}
\end{table}

\section{Experimental Setup}
\label{sec:setup}
In preparation to conducting our experiments with linguistically motivated tag
set modifications, a concrete setup for the experiments needed to be
established, which is presented in the following.

% fixme: tror nok dette bør flyttes og kortes ned. kan bakes inn med det som
% står under "Tag Set Mapping".
Our approach is somewhat semi-automatic in that we employ a "hybrid"
combination of manual annotation and automatic labeling in our experiments. Our
initial tags are gold standard, and using linguistic insight coupled with
computational considerations, we introduce new gold standard distinctions to
the tag set. The data sets with these tags are run through a pipeline with
automatically assigned tags in both training and testing, where the most
promising modifications are used "further". Hence, the tag set modifications
are not deterministic, nor do we try all possible combination; we only
experiment with linguistically motivated distinctions. We are in reality
altering the gold standard tags by the addition of more linguistic information.

As we want to investigate how we can improve the linguistic quality of a tag
set, we will only consider tag set modifications we deem linguistically
sensible.

\paragraph{Data Set Split}
As there was no existing standardized data set split of NDT due to its recent
development, we first needed to define separate sections for training, development and
testing. Our proposed sectioning of the treebank follows a 
standard 80-10-10 split. In establishing the  
split, care has been taken to preserve contiguous texts in the various 
sections while also keeping them balanced in terms of genre.

Our defined split into sections for training, development testing, and held-out
testing will be distributed with the future releases of the treebank, but 
details for replicating the split will also be made available at a companion website
for the final version of the paper. 
% Our proposed split has already been used in the Norwegian contribution to the
% Universal Dependencies project \cite{Ovr:Hoh:16}.

\paragraph{Tagger}
As our experiments during development required many repeated cycles of training and
testing for the different modified tag sets, we sought a PoS tagger that is both
reasonably fast and accurate. There is a often a considerable trade-off between
the two factors, as the most accurate taggers tend to suffer in terms of speed
due to their complexity. However, a tagger that achieves both close to
state-of-the-art accuracy as well as very high speed is TnT \cite{Bra:00}. TnT
was furthermore recently used to evaluate the recently proposed universal tag
set \cite{Pet:Das:McD:12}. The sum of these factors led to TnT being the tagger
of choice for our experiments.

\paragraph{Parser}
In choosing a syntactic parser for our experiments, we considered previous work
on dependency parsing of Norwegian, specifically that of
\newcite{Sol:Skj:Ovr:14}, who found the Mate parser \cite{Boh:10} to have  
the best performance for parsing of NDT. Furthermore, recent dependency
parser comparisons \cite{Cho:Tet:Ste:15} showed that Mate performed very well
on parsing of English, outperforming a range of contemporary state-of-the-art
parsers. We will be using Mate for gauging the effect of the tag set modifications in
our experiments.

\paragraph{Tag Set Mapping}
The tag set modifications are realized by mapping the existing tags to new,
more fine-grained tags "based" on associated morphological features to be
included in specified tags. The procedure maps the relevant existing tags to
new, more fine-grained tags including more relevant morphological features for
the applicable tokens, i.e., tokens that are assigned said tag and feature(s)
in the gold standard data. Hence, we are in actuality altering the gold
standard tags by the addition of more relevant linguistic information,
producing new gold standard annotations of higher quality.

%This is done by concatenating the original tag with the
%specified features for all applicable tokens

%contains the set of features in its morphological features.

\paragraph{Evaluation}
To evaluate tagging and parsing with the various tag set modifications in our
experiments, we employed commonly used evaluation metrics. Tagging is evaluated
in terms of accuracy (reported by the TnT-included \texttt{tnt-diff} script;
denoted \emph{Acc} in the following tables), while labeled attachment score
(LAS) and unlabeled attachment score (UAS) are computed by the
\texttt{eval.pl}\footnote{\url{http://ilk.uvt.nl/conll/software.html}} script
(used in the CoNLL shared tasks) to evaluate the parsing.

\paragraph{Baseline}
As a point of reference for the PoS tagging we will be reporting the commonly used
most-frequent-tag baseline, henceforth MFT.
This involves labeling each word with the tag it was
assigned most frequently in the training data. All unknown words, i.e., words not
seen in the training data, are assigned the tag most frequently assigned to
words seen only once.  The rationale is that we might expect unknown
and infrequent words to have similar properties given that they both occur rarely.

\paragraph{Tags \& Features} % <-- fixme: rename? 
As we seek to quantify the effects of PoS tagging in a realistic setting, i.e.,
in application to raw text, we choose to evaluate the parser on automatically
assigned PoS tags. For training, however, we have two
options: using either gold standard or automatically assigned tags. In order to
settle on a configuration, we conducted experiments with training on both gold
standard tags and
automatically assigned tags to see how the performance differs.
The results of these experiments, shown in Table \ref{tagconfigeval}, reveal that
the combination of training and testing on automatic tags is clearly superior
to training on gold standard tags and testing on automatic tags. Consequently,
the parser is both trained and tested on automatically assigned tags in the
remainder of our experiments.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Training} & \textbf{Testing} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        Gold & Gold & 90.15\% & 92.51\% \\
        Gold & Auto & 85.68\% & 88.98\% \\
        Auto & Auto & 87.01\% & 90.19\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of parsing with Mate using the various tag configurations,
        either using gold standard tags or automatically assigned from TnT.
        \emph{Gold} denotes gold standard tags, \emph{Auto} denotes
        automatically assigned tags from TnT.}
    \label{tagconfigeval}
\end{table}

Though some parsers make use of morphological features, % fixme: status for mate?
we removed all morphological features beyond the PoS tags in order to simulate a realistic setting.
Moreover, it is crucial that we remove these features when working with
automatically assigned tags, as such features would be provided by the gold
standard.

%For instance, if a verb token is erroneously tagged as a noun, we could potentially
%have a noun token with verbal features such as tense, which markedly obfuscates
%the training and parsing. Another important aspect is that we want to isolate
%the effect of PoS tags, necessitating the exclusion of morphological features.

In the following, we report on a set of experiments which evaluate increasingly
fine-grained PoS tag sets. The PoS tag sets are motivated by morphosyntactic
properties of Norwegian and are evaluated in terms of both tagger and parser
accuracy.

\section{Tag Set Optimization}
\label{sec:optimization}
By introducing more fine-grained linguistically motivated distinctions in a tag
set, we increase the linguistic information represented in the tags, which may
assist the parser in recognizing and generalizing syntactic patterns. However,
the addition of more linguistic information to the tags and thus a more
fine-grained tag set will most likely lead to a drop in tagger accuracy due to
the increase in complexity. The best tagging does not necessarily lead to the
best parse, hence it is interesting to investigate how linguistically informed
tag set modifications may affect the interplay between tagging and parsing.
%as
%we want to investigate how we can improve the linguistic quality of a tag set,
%we will
%only consider tag set modifications we deem linguistically sensible.
%we will not introduce tag set modifications which are not linguistically
%motivated. we will only consider distinctions we deem linguistically sensible.

\subsection{Baseline Experiments}
In an initial round of experiments, we concatenated the tag of each token with
its set of morphological features in order to map the original tag set to a
new, more fine-grained tag set (hereafter referred to as the \emph{full} tag
set). This resulted in a total of 368 tags, which is clearly very fine-grained.
The two initial tag sets, i.e., the original tag set comprising 19 tags and the
full tag set comprising 368 tags, thus represent two extremes in terms of
granularity. To evaluate the initial tag sets and investigate how the tag set
granularity would affect the performance of tagging and parsing, we ran
baseline experiments with the two tag sets. In Table \ref{inittagseteval}, we
report the results of these experiments. We see that the tagger accuracy
drastically drops when going from the original to the full tag set. The MFT
baseline for the original tag set is 94.14\%, while it drops by almost 9
percentage points to 85.15\% for the full tag set. TnT reports an accuracy of
97.47\% on the original tag set, which is reduced to 93.48\% for the full tag
set. These results confirm our hypothesis that the very high linguistic quality
in the full, fine-grained tag set comes at the expense of drops in tagger
performance.  However, the additional linguistic information provided by the
full tag set improves the parser performance. With the original tag set, Mate
reports an LAS of 87.01\% and a UAS of 90.19\%, which increases to 87.15\% and
90.39\%, respectively, using the full tag set. As we are looking for
linguistically informed distinctions that improve the syntactic parsing, these
results are promising and serve to indicate that additional linguistic
information assists the syntactic parsers, motivating the optimization of the
existing PoS tag set.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}lllll@{}}
        \toprule
        \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
        \textbf{LAS} & \textbf{UAS} \\
        \midrule
        Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        Full & 85.15\% & 93.48\% & \textbf{87.15\%} & \textbf{90.39\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the two initial tag sets.
        \emph{Acc} denotes tagger accuracy.}
    \label{inittagseteval}
\end{table}

\subsection{Tag Set Experiments}
We modified the tags for nouns (\texttt{subst}), verbs (\texttt{verb}),
adjectives (\texttt{adj}), determiners (\texttt{det}) and pronouns
(\texttt{pron}) in NDT by appending selected sets of morphological features to
each tag in order to increase the linguistic information expressed by the tags.
For each tag, we first experimented with each of the available features in
isolation before employing various combinations of them. We based our choices
of combinations on how promising the features are and what we deem worth
investigating in terms of linguistic utility, in order to see how the features
might interact.

The morphological properties of the various parts-of-speech are reflected in
the morphological features associated with the respective PoS tags. For
instance, as nouns in Norwegian can take on gender, definiteness and number,
additionally marked with genitive case, the treebank operates with features for
gender, definiteness, number and case.
%with "accompanying" values
In addition to morphological properties such as definiteness, tense and number,
all classes except for verbs have a \emph{type} feature that provides
information about the subtype, e.g., common nouns and proper nouns, while
pronouns can be reflexive, reciprocal, personal or interrogative.

%"Nearly" all tag set modifications leads to a drop in tagger accuracy, as
%expected. Few exceptions... We observed increases in parser accuracy scores
%with linguistically motivated tag set modifications for all PoS tags we
%experimented with, the most improved being noun with an increase in LAS from
%the original tag set by 1.80 percentage points.

\paragraph{Nouns}
In Norwegian, nouns are assigned gender (feminine, masculine or neuter),
definiteness (definite or indefinite) and number (singular or plural). There
is agreement in gender, definiteness and number between nouns and their
modifiers (adjectives and determiners). Additionally, NDT has a separate case
feature for distinguishing nouns in genitive case.

As nouns constitute the largest class in the treebank by far, the effects are
greatest for noun tokens in terms of overall change in performance. The results
of experimenting with tag set modifications for nouns are shown in Table
\ref{substresults}, and reveal that, apart from case, none of the tag set
modifications improves the tagging. However, they all give rise to increases in
parser accuracy scores. Genitive case marks possession, hence nouns marked with
genitive case are quite different from other nouns, taking a noun phrase as
complement. Distinguishing on type is useful and informative, as evident by the
presence of separate tags for proper and common nouns in many tag sets, such as
those of Penn Treebank and Stockholm-Umeå corpus.

The most informative features are definiteness, which
leads to an increase in LAS by 1.26 percentage points to 88.27\%, and type,
yielding an LAS of 88.07\%. Turning to combinations of features, we found that
the combination of type and case, as well as type and definiteness, were the
most promising, which led us to combine type, case and definiteness in a final
experiment, resulting in LAS of 88.81\% and UAS of 91.73\%, constituting large
increases from parsing with the original tag set, by 1.80 percentage points
and 1.54 percentage points, respectively.

\begin{table}
    \centering
    \smaller[1]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS}
        \\
        \midrule
        --- & 97.47\% & 87.01\% & 90.19\% \\
        Case & \textbf{97.48\%} & 87.63\% & 90.72\% \\
        Definiteness & 97.00\% & 88.27\% & 91.42\% \\
        Gender & 96.09\% & 87.21\% & 90.36\% \\
        Number & 96.37\% & 87.97\% & 91.00\% \\
        Type & 96.92\% & 88.07\% & 91.11\% \\
        Case \& definiteness & 97.03\% & 88.39\% & 91.44\% \\
        Type \& case & 96.92\% & 88.46\% & 91.51\% \\
        Type \& definiteness & 96.99\% & 88.44\% & 91.48\% \\
        %Type \& number & 96.37\% & 87.95\% & 90.95\% \\
        Type, case \& definiteness & 97.05\% & \textbf{88.81\%} &
        \textbf{91.73\%} \\
        %Type, definiteness \& number & 96.46\% & 88.06\% & 91.12\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for nouns.}
    \label{substresults}
\end{table}

\paragraph{Verbs}
Verbs are inflected for tense (infinitive, present, preterite, past perfect) in
Norwegian and additionally exhibit mood (imperative, indicative) and voice
(active or passive). Note that both voice and mood have only a single value in
the treebank, \texttt{pass} (passive) and \texttt{imp} (imperative),
respectively.  Verbs which are not passive are implicitly active, and verbs
which are not imperative are indicative.

Table \ref{verbresults} presents the results from tagging and parsing with
modified verb tags. Introducing features in isolation, mood is the only feature
leading to an increase in LAS, with a reported LAS of 87.04\%. Imperative
clauses are fundamentally different from indicative clauses, as they lack an
overt subject.  Although voice is a very distinguishing property for verbs, and
passive verbs appear in clauses very different from those with active verbs,
introducing this distinction in the tag set leads to a drop in LAS of 0.05
percentage points.  Distinguishing between the various tenses yields an LAS of
86.97\%. Combining the two most promising features, mood and tense, resulted in
an LAS of 87.12\% and UAS of 90.31\%.

In an additional experiment, we mapped the verb tenses (and mood, in the case
of imperative) to finiteness. All verbs have finiteness, hence this distinction
has broad coverage. This mapping is syntactically grounded as finite verbs and
nonfinite verbs appear in completely different syntactic constructions, and
proved to greatly improve the parsing, as we saw the overall largest parser
accuracy scores, with 87.30\% for LAS and 90.43\% for UAS, 0.29 and 0.24
percentage points higher than the baseline, respectively. This coincides with
the observations seen for Swedish in \newcite{Ovr:08}, where finiteness was
found to be a very beneficial linguistic feature for parsing.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        Mood & 97.43\% & 87.04\% & 90.19\% \\
        Tense & 97.30\% & 86.97\% & 90.18\% \\
        Voice & 97.45\% & 86.96\% & 90.09\% \\
        Mood \& tense & 97.31\% & 87.12\% & 90.31\% \\
        Voice \& tense & 97.28\% & 86.99\% & 90.15\% \\
        Mood, tense \& voice & 97.27\% & 86.83\% & 90.05\% \\
        Finiteness & 97.35\% & \textbf{87.30\%} & \textbf{90.43\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for verbs.}
    \label{verbresults}
\end{table}

\paragraph{Adjectives}
Adjectives agree with the noun they modify in terms of gender, number and
definiteness in Norwegian. Adjectives are also inflected for degree (positive,
comparative or superlative).

In Table \ref{adjresults}, we report the results of modifying the \texttt{pron}
tag in NDT. All features except for number led to increases in parser accuracy
scores, the most successful of which is degree with a reported LAS of 87.29\%,
while distinguishing adjectives on definiteness yield an LAS of 87.14\% and
gender leads to LAS of 87.10\%.

%Combining the most promising features, we
%found that none of the combinations surpass the LAS with the distinction of
%degree alone. Definiteness and degree with an LAS of 87.23\% and definiteness
%and number with an LAS of 87.27\% and UAS identical to that of degree alone.

Turning to combinations of features, definiteness and number achieve the best
results, very close to that of degree, with 0.02 percentage points lower LAS
and identical UAS. Adjectives agree with their head noun and determiner in
definiteness and number, making this an expected improvement. The combination
of definiteness and degree is also quite promising, obtaining LAS of 87.23\%
and UAS of 90.39\%. It is interesting that none of the combinations surpass the
experiment with degree alone, which indicates that degree does not interact
with the other features in any syntactically significant way.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS}
        \\
        \midrule
        --- & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        Definiteness & 96.84\% & 87.14\% & 90.29\% \\
        Degree & 97.41\% & \textbf{87.29\%} & \textbf{90.44\%} \\
        Gender & 96.89\% & 87.10\% & 90.25\% \\
        Number & 96.71\% & 86.99\% & 90.10\% \\
        Type & 97.40\% & 87.11\% & 90.25\% \\
        Definiteness \& degree & 96.81\% & 87.23\% & 90.39\% \\
        Definiteness \& gender & 96.31\% & 87.18\% & 90.39\% \\
        Definiteness \& number & 96.78\% & 87.27\% & \textbf{90.44\%} \\
        %Degree \& gender & 96.87\% & 87.00\% & 90.16\% \\
        %Degree \& number & 96.76\% & 87.13\% & 90.26\% \\
        %Definiteness, degree \& number & 96.81\% & 87.14\% & 90.30\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for adjectives.}
    \label{adjresults}
\end{table}

\paragraph{Determiners}
Like adjectives, determiners in Norwegian agree with the noun they modify in
terms of gender, number and definiteness. Possessive pronouns are
treated/analyzed as determiners in NDT.

%For a number of reasons, we do not report results from combinations of features
%for determiners, as the features could not be combined in any meaningful way.
%For instance, determiners in plural never have marked definiteness, hence this
%combination would only apply to determiners in singular and roughly correspond
%to the distinction on definiteness alone. Moreover, determiners in plural are
%not distinguished in terms of gender (thus ruling out the combination of number
%and gender), neither are definite determiners. Another aspect taken into
%consideration is that the most promising distinction, definiteness, applies to
%such a small number of tokens that more fine-grained distinctions would be
%overly sparse.

The results from the experiments with determiners are shown in Table
\ref{detresults}.  Introducing the distinction of type led to an increase in
tagger accuracy by 0.14 percentage points to 97.61\%, while marginally
impacting the parsing, with LAS of 87.00\%, 0.01 percentage points below that
of the original tag set.  The increase in tagger accuracy when introducing the
distinction of type is noteworthy, as we expected the finer granularity to lead
to a decrease in accuracy. This serves to indicate that more fine-grained
distinctions for determiners, which is a quite disparate category in the
treebank, may be quite useful for tagging. However, as it has negative impact
on the syntactic parsing, we can conclude that the type of a determiner does
not assist in generalizing syntactic patterns, as most determiners, regardless
of type, appear in the same syntactic constructions (i.e., before a noun or
noun phrase).

Gender, on the other hand, improved the parsing, but complicated the tagging,
as the various genders are often difficult to differentiate, especially so in
the case of masculine and feminine, which share many of the same determiners.
The number of a determiner, i.e., singular or plural, led to a small increase
in tagger accuracy and LAS, while marginally lower UAS, 90.18\%, 0.01
percentage points lower than that of the original tag set. The
introduction of definiteness to the determiners led to the best parsing
results, LAS of 87.30\% and UAS of 90.42\%, while also increasing the tagger
accuracy slightly. The increase in LAS and UAS is rather interesting, as there
are only 121 determiner tokens with marked definiteness in the development
data. As this accounts for a very small number of tokens, we did not consider
further fine-grained modifications with definiteness. This goes to show that
tokens with overt definiteness have noticeable impact on the syntactic parsing,
and that distinguishing on definiteness is very beneficial.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        --- & 97.47\% & 87.01\% & 90.19\% \\
        Definiteness & 97.49\% & \textbf{87.30\%} & \textbf{90.42\%} \\
        Gender & 97.28\% & 87.09\% & 90.31\% \\
        Number & 97.49\% & 87.04\% & 90.18\% \\
        Type & \textbf{97.61\%} & 87.00\% & 90.11\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for determiners.}
    \label{detresults}
\end{table}

\paragraph{Pronouns}
Pronouns in Norwegian include personal, reciprocal, reflexive and
interrogative. They can exhibit gender, number and person. Personal pronouns
have case (accusative or nominative).

The results in Table \ref{pronresults} show that number, person and type are
the most informative features for parsing, with LAS of 87.21\%, 87.22\% and
87.19\%, respectively. However, when combining number and person, we observe a
drop by more than 0.2 percentage points, indicating that these features do not
interact in any syntactically distinctive way. The most interesting observation
is that all experiments exceed the tagging accuracy of the original tag set,
the most improved being the most fine-grained distinction, namely type, case
and number combined, obtaining a tagger accuracy of 97.52\%. This shows that
the introduction of more fine-grained distinctions for pronouns is beneficial
and aids the PoS tagger in disambiguating ambiguous words. While case alone
yields an LAS of 87.08\%, we found that the combination of type and case, which
is the most successful experiment in terms of parser performance, yields the
second highest tagging accuracy of 97.51\%. The reason for this is that
pronouns of different type and personal pronouns of different case exhibit
quite different properties and appear in different constructions. Pronouns in
nominative case (i.e., subjects) primarily occur before the main verb, while
pronouns in accusative case (i.e., objects) occur after the main verb, as
Norwegian exhibits so-called V2 word order, requiring that the finite verb of a
declarative clause appears in the second position, hence its name. The
combination of type and number comes in close to the performance of type and
case, with LAS of 87.27\% and UAS identical to that of type and case.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llll@{}}
        \toprule
        \textbf{Feature(s)} & \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        --- & 97.47\% & 87.01\% & 90.19\% \\
        Case & 97.50\% & 87.08\% & 90.21\% \\
        Gender & 97.48\% & 87.06\% & 90.23\% \\
        Number & 97.49\% & 87.21\% & 90.33\% \\
        Person & 97.49\% & 87.22\% & 90.32\% \\
        Type & 97.48\% & 87.19\% & 90.40\% \\
        Number \& person & 97.49\% & 96.98\% & 90.16\% \\
        Type \& case & 97.51\% & \textbf{87.30\%} & \textbf{90.41\%} \\
        Type \& number & 97.49\% & 87.27\% & \textbf{90.41\%} \\
        Type \& person & 97.49\% & 87.00\% & 90.14\% \\
        Type, case \& number & \textbf{97.52\%} & 87.11\% & 90.36\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of experiments with modified PoS tags for pronouns.}
    \label{pronresults}
\end{table}

\subsection{Optimized Tag Set}
\begin{table*}[ht]
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llllll@{}}
        \toprule
        \textbf{Category} & \textbf{Feature(s)} & \textbf{MFT} &
        \textbf{Acc} & \textbf{LAS} & \textbf{UAS} \\
        \midrule
        \emph{Original} & --- & 94.14\% & 97.47\% & 87.01\% & 90.19\% \\
        Noun & Type, case \& definiteness & 89.61\% & 97.05\% & 88.81\% &
        91.73\% \\
        Verb & Finiteness & 93.72\% & 97.35\% & 87.30\% & 90.43\% \\
        Adjective & Degree & 94.13\% & 97.41\% & 87.29\% & 90.44\% \\
        Determiner & Definiteness & 94.13\% & 97.49\% & 87.30\% & 90.42\% \\
        Pronoun & Type \& case & 94.12\% & 97.51\% & 87.30\% & 90.41\% \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the most successful tag set
        modification for each category.}
    \label{respectiveresults}
\end{table*}

The most successful tag set modification for each PoS and their results are
seen in Table \ref{respectiveresults}. Nouns benefit by far the most from the
introduction of more fine-grained linguistically motivated distinctions, with
an LAS of 88.81\% and UAS of 91.73\%. We observe that the most promising tag
set modifications for verbs, adjectives, determiners and pronouns all reach LAS
of \textasciitilde 87.30\% and UAS of \textasciitilde 90.40\%. To investigate
the overall effect of these tag set modifications, we tested each of the
improvements in parser accuracy scores from baseline for statistical
significance using Dan Bikel's randomized parsing evaluation comparator
script\footnote{Available as \texttt{compare.pl} at
    \url{http://ilk.uvt.nl/conll/software.html}}, as used in the CoNLL shared
tasks. For the most successful tag set modification for each of the categories
seen in Table \ref{respectiveresults}, the difference in LAS from the original
tag set is statistically significant at significance level 0.05 (\emph{p}-value
< 0.05), as are all differences in UAS, except for verbs with finiteness
(\emph{p}-value 0.15) and pronouns with type and case (\emph{p}-value 0.06).

An overview of the tags in the optimized tag set can be seen in Table
\ref{optimizedtagset}, comprising three new tags for adjectives, two for
determiners, six for pronouns, seven for nouns and two for verbs, totaling 20
tags. Appending these to the original tag set comprising 19 tags, we reach a
total of 39 tags for NDT.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Tag} & \textbf{Description} \\
        \midrule
        \texttt{adj|komp} & Comparative adjective \\
        \texttt{adj|pos} & Positive adjective \\
        \texttt{adj|sup} & Superlative adjective \\
        \texttt{det|be} & Definite determiner \\
        \texttt{det|ub} & Indefinite determiner \\
        \texttt{pron|pers} & Personal pronoun \\
        \texttt{pron|pers|akk} & Personal pronoun, accusative \\
        \texttt{pron|pers|nom} & Personal pronoun, nominative \\
        \texttt{pron|refl} & Reflexive pronoun \\
        \texttt{pron|res} & Reciprocal pronoun \\
        \texttt{pron|sp} & Interrogative pronoun \\
        \texttt{subst|appell} & Common noun \\
        \texttt{subst|appell|be} & Common noun, def. \\
        \texttt{subst|appell|be|gen} & Common noun, def., genitive \\
        \texttt{subst|appell|ub} & Common noun, indef. \\
        \texttt{subst|appell|ub|gen} & Common noun, indef., genitive \\
        \texttt{subst|prop} & Proper noun \\
        \texttt{subst|prop|gen} & Proper noun, genitive \\
        \texttt{verb|fin} & Finite verb \\
        \texttt{verb|infin} & Nonfinite verb \\
        \bottomrule
    \end{tabular}
    \caption{Overview of the optimized tag set.}
    \label{optimizedtagset}
\end{table}

\paragraph{Final Evaluation}
In Table \ref{finalresults}, we show the results of parsing with the optimized
tag set on the held-out test set, compared to the results obtained with the
original tag set. We see significant improvements from the original tag set on
both the development data and the held-out test data set. The improvement in
LAS on the development data is 1.86 percentage points, while 1.91 percentage
points on the held-out test data. These results indicate that the additional
linguistic information in the tags of our optimized tag set greatly aids
syntactic parsing and that optimizing an existing PoS tag set for a downstream
application can be useful and beneficial.

\begin{table}
    \centering
    \smaller[0.5]
    \begin{tabular}{@{}llllll@{}}
        \toprule
        \textbf{Data} & \textbf{Tag Set} & \textbf{MFT} & \textbf{Acc} &
        \textbf{LAS} & \textbf{UAS} \\
        \midrule
        \multirow{2}{*}{Dev}
        & Original & \textbf{94.14\%} & \textbf{97.47\%} & 87.01\% & 90.19\% \\
        %Full & 85.15\% & 93.48\% & 87.15\% & 90.39\% \\
        & Optimized & 85.15\% & 96.85\% & \textbf{88.87\%} & \textbf{91.78\%} \\
        \midrule
        \multirow{2}{*}{Test}
        & Original & \textbf{94.22\%} & \textbf{97.30\%} & 86.64\% & 90.07\% \\
        %& Full & 84.27\% & 92.79\% & 87.03\% & 90.16\% \\
        & Optimized & 88.08\% & 96.35\% & \textbf{88.55\%} & \textbf{91.41\%} \\
        \bottomrule
    \end{tabular}
    \caption{Results of tagging and parsing with the optimized tag set,
        compared to the initial tag sets. Trained and tested using
        automatically assigned tags from TnT.}
    \label{finalresults}
\end{table}

%\section{Optimized Pipeline}
%\label{sec:pipeline}

\section{Summary and Future Work}
\label{sec:summary}
The improvements in parser performance with our optimized tag set indicate that
a more fine-grained PoS tag set may assist syntactic parsers in recognizing and
generalizing syntactic patterns, while potentially compromising the performance
of PoS taggers.

There are several aspects of this thesis that can be further explored in future
work, including extrinsic evaluation of the effects of PoS tag sets on other
downstream NLP applications besides parsing, such as sentiment analysis and
named entity recognition. These applications often require tagged data, but are
markedly different from syntactic parsing, hence the evaluation would involve
investigating an entirely different aspect of the effects of tag set
granularity.

% \section*{Acknowledgments}

\bibliographystyle{acl2016}
\bibliography{references}
% \appendix
%\section{Supplemental Material}
%\label{sec:supplemental}

\end{document}